---
title: ""
output:
  pdf_document:
    number_sections: TRUE
geometry: "left = 2.5cm, right = 2cm, top = 2cm, bottom = 2cm"
fontsize: 11pt
header-includes:
  - \usepackage{float}
  - \usepackage{sectsty}
  - \usepackage{paralist}
  - \usepackage{setspace}\spacing{1.25}
  - \usepackage{fancyhdr}
  - \usepackage{lastpage}
  - \usepackage{dcolumn}
  - \usepackage{pdflscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}

---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(include = FALSE,fig.lp = '')
options(tinytex.verbose = TRUE)
```



```{=tex}
\allsectionsfont{\centering}
\subsectionfont{\raggedright}
\subsubsectionfont{\raggedright}
```
\pagenumbering{gobble}


\begin{centering}

\vspace{3cm}

\Large

\doublespacing
{\bf ASSESSMENT 2: DATA ANALYSIS \\Part B: Report}

\vspace{1cm}


\includegraphics[width=.4\textwidth]{UTSlogo.png}

\Large

\doublespacing
{\bf 36103 Statistical Thinking for Data Science\\University Of Technology Sydney}

\vspace{1 cm}

\normalsize
\singlespacing

\Large{\textbf{Group Name}}\\
\doublespacing
Retail Group\\

\vspace{1 cm}

\Large{\textbf{Group Members}}\\
\doublespacing
Leah Nguyen\\
Tony Tan\\
Anmol Mittal\\
Ben McKinnon\\
Kasun Caldera\\
Manasa Burli Nagendra\\

\vspace*{1\baselineskip}

\Large{\textbf{Date}}\\
\doublespacing
October, 2021

\end{centering}

\pagenumbering{roman}

\newpage

<!---Table of Content (TOC)-->
\centering
\raggedright
\newpage
\tableofcontents
\newpage

<!---Sections-->
<!---Executive Summary-->
```{r intro, child = 'executive.Rmd'}
```
\pagenumbering{arabic}

<!---Introduction-->
```{r intro, child = 'introduction.Rmd'}
```


<!---Data-->
```{r data, child='data.Rmd'}
```


<!---EDA-->
```{r data, child='eda.Rmd'}
```


<!---Method-->
```{r method, child = 'method.Rmd'}
```


<!---Limitations-->
```{r limitations, child = 'limitations.Rmd'}
```

\newpage
<!---References-->
```{r biblio, child = 'bibliography.Rmd'}
```

\newpage

<!---Appendix-->
\blandscape
```{r appendix, child = 'appendix1.Rmd'}
```
\elandscape

```{r appendix, child = 'appendix2.Rmd'}
```

```{r appendix, child = 'appendix3.Rmd'}
```